---
title: "Dow Jones Industrial"
output: html_notebook
Auther: Sakshi Patil
editor_options: 
  markdown: 
    wrap: 72
---

```{r}
install.packages('quantmod')
install.packages('tseries')
install.packages('forecast')
install.packages('car')
install.packages('FinTS')
install.packages('rugarch')
install.packages('utf8')
install.packages('ggplot2')
library(quantmod)
library(tseries)
library(car)
library(FinTS)
library(utf8)
library(ggplot2)
library(forecast)
library(rugarch)
lapply(packages, require, character.only = TRUE) 
```

```{r}
packages = c('quantmod','car','forecast','tseries','FinTS', 'rugarch','utf8','ggplot2')
install.packages(packages, dependencies = TRUE)
lapply(packages, require, character.only=TRUE)
```

**Downloading the stock data**

```{r}
getSymbols(Symbols = '^DJI', 
           src = 'yahoo', 
           from = as.Date('2016-01-01'), 
           to = as.Date('2023-12-31'),
           periodicity = 'daily')

Stock_price = na.omit(DJI$DJI.Adjusted) # Adjusted Closing Price, omitting null values
class(Stock_price) # xts (Time-Series) Object
View(Stock_price)

Stock_return = na.omit(diff(log(Stock_price))); plot(Stock_return) # Returns of TATA
plot(Stock_price)

```

**ADF Test for Stationarity**

**Objective:**

To check whether the stock return data and closing price data are
stationary or not

**Analysis:**

The Augmented Dickey-Fuller Test is a statistical test used to determine
whether a given time series data is stationary or not. Stationarity of
time series data indicates that the mean of the series do not change
over time, making it easier to model and forecast.

H0: Time Series is not stationary H1: Times Series is stationary

Here p-value for closing-price:0.9758 > 0.05, indicating series not
stationary

Whereas, p-values for return series is: 0.01 < 0.05 hence series is
stationary

**Results:**

p-values for DJI returns series is: 0.01 < 0.05 hence
'Stock_return' series is stationary.

**Managerial Implication:**

With stationary data, modeling and forecasting stock returns become
more reliable as the statistical properties of the data remain
consistent over time.

```{r}
# Augmented Dickey-Fuller (ADF) Test for stationarity
adf_test_wn = adf.test(Stock_price); adf_test_wn # for price
adf_test_wn = adf.test(Stock_return); adf_test_wn # for return

```

**Ljung Box test for autocorrelation**

**Objective:**

To check whether the stock return data series has autocorrelation or not

**Analysis:**

The Box-Pierce test assesses the presence of autocorrelation in a time
series model. In this case, the test was applied to both returns of Dow
Jones Industrial Avg data

Null Hypothesis (H0): There is no autocorrelation in the time series
Alternate Hypothesis (H1): There is autocorrelation in the time series

Here p-value for return data series :3.619e-13 \< 0.05, indicating
series has autocorrelation

**Results:**

p-values for Dow Jones Industrial Avg data returns series is: 3.619e-13
< 0.05 hence 'Stock_return' series has autocorrelation.

**Managerial Implication:**

since data is auto-correlated, we can go for time-series modelling of
data instead of regression

```{r}
#Ljung Box test for autocorrelation
lb_test_wn = Box.test(Stock_return); lb_test_wn

```

**ACF and PACF Plot**

**Objective:** ACF: p and PACF: q

To assess the presence of autocorrelation in a time series data
set(correlation between a time series and its lagged versions). In other
words, it indicates whether the past values of a series can be used to
predict its future values

**Analysis:**

1.  The ACF plots for the series "Stock_Return" shows the
    autocorrelation coefficients for different lags. The autocorrelation
    coefficient at lag k measures the linear dependence between the
    series and its lagged version by k periods. In this specific plot,
    the autocorrelation coefficients are plotted on the y-axis, and the
    lags are plotted on the x-axis.
2.  The PACF plot specifically focuses on identifying the lags at which
    the series exhibits significant autocorrelations, net of the
    correlations at shorter lags.

**Results:**

1.  From the ACF plot, most of the autocorrelation coefficients fall
    within the shaded area around zero. This suggests that there is
    *slightly* statistically significant autocorrelation present in the
    series at any of the lags depicted.
2.  In other words, the past values of the stock returns slightly appear
    to have a significant predictive power for future returns.
3.  There might be a significant partial autocorrelation at a specific
    lag. This suggests that the series exhibits an autocorrelation at
    that lag, which is not explained by the correlations at shorter
    lags.

**Managerial Implication:**

The presence of a significant partial autocorrelation in stock returns
suggests that past values of the stock returns might have a predictive
power for future returns

```{r}
#ACF and PACF plots
acf(Stock_return)
pacf(Stock_return)
```

**Manual ARIMA Model**

**Objective:**

To interpret the Auto Regressive(AR) and Moving Average(MA) order values
for ARMA Modelling forecast using manual ARMA

**Analysis:**

1.  The ARIMA model is a time series forecasting method that combines
    auto-regressive (AR), differencing (I), and moving average (MA)
    components.
2.  The method was used for different combinations of coefficients,
    output shows the coefficients estimated for the ARIMA(1,0,5) model
    has least sigma-square.

**Results:**

Out of all the orders tested for ARIMA order, order with (1,0,5) has the
least variance (sigma-square):

Model Coefficients: AR (autoregressive) Coefficient (ar1): -0.8003 MA
(moving average) Coefficients (ma1-ma5): 0.6925, 0.0014, 0.0919,
-0.0645, 0.0457 Intercept Coefficient: 4e-04 Estimated Variance
(sigma^2): 0.000126

**Managerial Implication:**

The fitted ARIMA(1,0,5) model provides valuable insights into the
dynamics of the stock returns data and offers a framework for
forecasting future returns.

```{r}
# ARIMA Model
# ARIMA (1, 0, 0) or AR(1)
ar1 = arima(Stock_return, order = c(1, 0, 0)); ar1

# ARIMA (2, 0, 0) or AR(2)
ar2 = arima(Stock_return, order = c(2, 0, 0)); ar2

# ARIMA (0, 0 , 1) or MA(1)
ma1 = arima(Stock_return, order = c(0, 0, 1)); ma1

# ARIMA (0, 0, 2) or MA(2)
ma2 = arima(Stock_return, order = c(0, 0, 2)); ma2

# ARIMA (0, 0, 3) or MA(3)
ma3 = arima(Stock_return, order = c(0, 0, 3)); ma3

# ARIMA (0, 0, 4) or MA(4)
ma4 = arima(Stock_return, order = c(0, 0, 4)); ma4

# ARIMA (1, 0, 1) or ARMA(1, 1)
arma11 = arima(Stock_return, order = c(1, 0, 1)); arma11

# ARIMA (1, 0, 2) or ARMA(1, 2)
arma12 = arima(Stock_return, order = c(1, 0, 2)); arma12

# ARIMA (1, 0, 3) or ARMA(1, 3)
arma13 = arima(Stock_return, order = c(1, 0, 3)); arma13

# ARIMA (1, 0, 4) or ARMA(1, 4)
arma14 = arima(Stock_return, order = c(1, 0, 4)); arma14

# ARIMA (1, 0, 5) or ARMA(1, 5)
arma15 = arima(Stock_return, order = c(1, 0, 5)); arma15
```

**Auto ARIMA Model**

**Objective:**

To interpret the Auto Regressive(AR) and Moving Average(MA) order values
for ARMA Modelling forecast using Auto-ARMA method

**Analysis:**

1.  The ARIMA model is a time series forecasting method that combines
    auto-regressive (AR), differencing (I), and moving average (MA)
    components.
2.  The method was used to find out best order os ARIMA for further
    ARIMA Modelling

**Results:**

Auto-ARIMA gives order of (1,0,5) as the best order which matches with
the manual ARIMA Model:

Model Coefficients: AR (autoregressive) Coefficient (ar1): -0.8003 MA
(moving average) Coefficients (ma1-ma5): 0.6925, 0.0014, 0.0919,
-0.0645, 0.0457 Mean: 4e-04 Variance (sigma\^2): 0.0001264

**Managerial Implication:**

The fitted ARIMA(1,0,5) model provides valuable insights into the
dynamics of the stock returns data and offers a framework for
forecasting future returns.

```{r}
# Auto ARIMA
arma_pq_jj_ds = auto.arima(Stock_return); arma_pq_jj_ds # chosen order is (1,0,5)
```

**Auto-Correlation test for residuals in ARIMA Modelling**

**Objective:**

To find out whether there is auto-correlation between residuals of ARIMA
model and hence define validity of using ARIMA Forecast

**Analysis:**

1.  The Box-Pierce test is used to assess the goodness-of-fit of a time
    series model by examining the autocorrelation of its residuals.
2.  In this analysis, the test was applied to the residuals of an ARMA
    model, specifically to evaluate whether there is significant
    autocorrelation present.

Null Hypothesis (H0): There is no autocorrelation in the residuals of
ARIMA Model Alternate Hypothesis (H1): There is autocorrelation in the
residuals of ARIMA Model

Here p-value for ARIMA Model residuals is :0.874 > 0.05, indicating
residuals have no auto-correlation

**Results:**

The obtained p-value of 0.874 is substantially greater than the
conventional significance level of 0.05. Therefore, there we do not
reject the null hypothesis and there is no autocorrelation in the
residuals.

**Managerial Implication:**

The Box-Pierce test results provide reassurance regarding the adequacy
of the ARMA model in forecasting the data, hence we can forecast the
data with ARIMA Modelling

```{r}
# Autocorrelation test for residuals
lb_test_arma_pq_jj_ds = Box.test(arma_pq_jj_ds$residuals); lb_test_arma_pq_jj_ds #no-autocorrelation (do not reject null) good ARIMA Model, now forecast
```

**Forecasting with ARIMA Modelling**

**Objective:**

To forecast the stock data returns using ARIMA Modelling

**Analysis:**

Since there is no auto-correlation between residuals of ARIMA(1,0,5), we
forecasted the stock values for next 500 time units.

**Results:**

ARIMA modelling forecasted the stock values for next 500 time units.

**Managerial Implication:**

Now that we have forecasted the stock values using ARIMA modelling we
test its validity.

```{r}
#ARIMA Modelling
jj_ds_fpq = forecast(arma_pq_jj_ds, h = 500)
plot(jj_ds_fpq)
```

**Auto-Correlation test for residual square**

**Objective:**

To predict the model's volatility based on auto-correlation between
residual-square(variance) of ARIMA model

**Analysis:**

1.  The Box-Pierce test is used to assess the goodness-of-fit of a time
    series model by examining the autocorrelation of its
    residuals-square.
2.  In this analysis, the test was applied to the residuals-square of an
    ARIMA model, specifically to evaluate whether there is significant
    autocorrelation present or not.

Null Hypothesis (H0): There is no autocorrelation in the
residuals-square of ARIMA Model Alternate Hypothesis (H1): There is
autocorrelation in the residuals-square of ARIMA Model

Here p-value for ARIMA Model residuals-square is : 2.2e-16 < 0.05,
indicating residuals have no auto-correlation

**Results:**

The obtained p-value of 2.2e-16 is substantially less than the
conventional significance level of 0.05. Therefore, we reject the null
hypothesis and there is autocorrelation in the residuals.

**Managerial Implication:**

The Box-Pierce test results provide information that there is
significant autocorrelation amongst residual-square and hence
significant volatility.

```{r}
# Autocorrelation test for residual square(to check autocorrelation in residual square)
lb_test_arma_pq_jj_ds = Box.test(jj_ds_fpq$residuals^2); lb_test_arma_pq_jj_ds #there is autocorrelation
```

**ARCH-LM test for heteroskedasticity**

**Objective:**

To predict the model's volatility based on ARCH-LM test for
heteroskedasticity in residual-square of ARIMA model

**Analysis:**

1.  The ARCH (Autoregressive Conditional Heteroscedasticity) LM-test is
    used to assess whether there are ARCH effects present in the
    residuals of a time series model.
2.  ARCH effects refer to the presence of conditional
    heteroscedasticity, where the variance of the residuals depends on
    past values of the series.

Null Hypothesis (H0): There is no ARCH effects(Homoskedastic) Alternate
Hypothesis (H1): There is ARCH effects present in
residual-square(variance) (Heteroskedastic)

Here p-value for ARIMA Model residuals-square is : 2.2e-16 < 0.05,
indicating heteroskedasticity in residual-square

**Results:**

The obtained p-value of 2.2e-16 is substantially less than the
conventional significance level of 0.05. Therefore, we reject the null
hypothesis and there is heteroskedasticity in residual-square.

**Managerial Implication:**

The significant ARCH effects detected in the squared residuals indicate
that the current time series model fails to account for conditional
heteroscedasticity adequately.

```{r}
#Test for Volatility Clustering or Heteroskedasticity: Box Test 

nse_ret_arch_test = ArchTest(jj_ds_fpq$residuals^2, lags = 20) # H0: No ARCH Effects, hence homoskedastic, go for GARCH
nse_ret_arch_test #here there is ARCH effect hence residual square is heteroskedastic

```

**GARCH Modelling for ARIMA model**

**Objective:**

To model ARIMA data with GARCH and test the fit of GARCH model with data

**Analysis:**

The GARCH tests the fit of the GARCH model to the data. These statistics
and tests evaluate the adequacy of the model in capturing the volatility
dynamics of the time series.

**Results:**

Joint Statistic: The joint statistic is 89.5651, indicating the overall
goodness-of-fit of the GARCH model. Individual Statistics: mu (Mean):
0.2169 omega (Constant Term in Volatility Equation): 25.7892 alpha1
(Coefficient of Lagged Squared Errors): 4.2219 beta1 (Coefficient of
Lagged Conditional Variance): 2.4185 Asymptotic Critical Values (10%,
5%, 1%): For Joint Statistic: 1.07, 1.24, 1.6 For Individual Statistics:
0.35, 0.47, 0.75

**Managerial Implication:**

GARCH model provides insights into the volatility dynamics of the time
series, the results of the goodness-of-fit test suggest that the model
may have limitations in accurately capturing the underlying patterns.

```{r}
#GARCH Model

garch_model1 = ugarchspec(variance.model = list(model = 'sGARCH', garchOrder = c(1,1)), mean.model = list(armaOrder = c(0,0), include.mean = TRUE)) # since we are modelling mean
nse_ret_garch1 = ugarchfit(garch_model1, data = jj_ds_fpq$residuals^2);
nse_ret_garch1
```

**ARCH-LM test for heteroskedasticity of GARCH Modelling**

**Objective:**

To predict the GARCH model's volatility based on ARCH-LM test for
heteroskedasticity in residual of GARCH model

**Analysis:**

1.  The ARCH (Autoregressive Conditional Heteroscedasticity) LM-test is
    used to assess whether there are ARCH effects present in the
    residuals of a time series model.
2.  ARCH effects refer to the presence of conditional
    heteroscedasticity, where the variance of the residuals depends on
    past values of the series.

Null Hypothesis (H0): There is no ARCH effects(Homoskedastic) Alternate
Hypothesis (H1): There is ARCH effects present in
residual-square(variance) (Heteroskedastic)

Here p-value for ARIMA Model residuals-square is : 2.2e-16 \< 0.05,
indicating heteroskedasticity in residual

**Results:**

The obtained p-value of 2.2e-16 is substantially less than the
conventional significance level of 0.05. Therefore, we reject the null
hypothesis and there is heteroskedasticity in residual of GARCH Model
indicating the inadquacy of GARCH Model.

**Managerial Implication:**

The significant ARCH effects detected in the residuals indicate that the
current GARCH model fails to account for conditional heteroscedasticity
adequately.

```{r}
#Heteroskedasticity test
# Test for Volatility Clustering or Heteroskedasticity: ARCH Test
gar_resd = residuals(nse_ret_garch1)
stk_ret_arch_test1 = ArchTest(gar_resd, lags = 1) # H0: No ARCH Effects
stk_ret_arch_test1 # Inference : Return Series is Heteroskedastic (Has Volatility Clustering
```

**GARCH Modelling**

**Objective:**

To test the fit of GARCH Model on stock return data

**Analysis:**

The GARCH tests the fit of the GARCH model to the data. These statistics
and tests evaluate the adequacy of the model in capturing the volatility
dynamics of the time series.

**Results:**

Conditional Variance Dynamics: GARCH Model: sGARCH(1,1) Mean Model:
ARFIMA(1,0,5) Distribution: Normal Optimal Parameters: Parameter
estimates and robust standard errors for AR (ar1), MA (ma1-ma5), omega,
alpha1, and beta1 are provided. Log-Likelihood and Information Criteria:
Log-likelihood: 6733.164 Information criteria including Akaike, Bayes,
Shibata, and Hannan-Quinn are reported. Adjusted Pearson Goodness-of-Fit
Test: Tests for goodness-of-fit of the model. Extremely low p-values
indicate significant discrepancies between the model and the data.

**Managerial Implication:**

Model Adequacy: The GARCH model with specified parameters captures the
volatility dynamics and mean behavior to some extent.

Model Evaluation: Diagnostic tests indicate no significant issues such
as serial correlation or ARCH effects in the residuals. However,
goodness-of-fit tests reveal discrepancies between the model and the
data.

Forecast Reliability: Caution should be exercised when using forecasts
from the GARCH model, considering potential limitations in the model
fit.

Risk Management: Inaccurate volatility modeling can impact risk
management decisions. Further evaluation and refinement of the GARCH
model or consideration of alternative models may be necessary for robust
risk assessment.

GARCH model provides insights into volatility dynamics, discrepancies in
goodness-of-fit tests suggest potential limitations.

```{r}
garch_modelf = ugarchspec(variance.model = list(model = 'sGARCH', garchOrder = c(1,1)), mean.model = list(armaOrder = c(1,5), include.mean = TRUE))
stk_ret_garch = ugarchfit(garch_modelf, data = Stock_return);stk_ret_garch
```

**GARCH Modelling**

**Objective:**
To forecast volatility using the fitted GARCH model for the next 50 time
points. 

**Analysis:**
Used the 'ugarchforecast' function to generate
volatility forecasts for the next 50 time points.

**Results:**

**GARCH Model Forecast:**
Model: sGARCH
- Horizon: 50
- Roll Steps: 0
- Out of Sample: 0

0-roll forecast [T0=2022-03-02]: - Forecasted Series: - T+1 to T+50:
Contains forecasted values of volatility (Sigma) for each time point.

**Implication:**
The forecasted values represent the predicted volatility for the next 50
time points based on the fitted GARCH model. These forecasts can be
useful for risk management and decision-making, providing insights into
the expected future volatility of the financial time series.

```{r}
# GARCH Forecast
stk_ret_garch_forecast1 = ugarchforecast(stk_ret_garch, n.ahead = 50); stk_ret_garch_forecast1
```

```{r}
plot(stk_ret_garch_forecast1)
```
